### 2024-09-23 paper/paper.qmd
Added onto data in the papaer, as well as the introduction, added replications of paper's graphs in relications

---
title: "My title"
subtitle: STA304 Paper 1 Submission
author: 
  - Veyasan Ragulan
  - Another author
thanks: "Code and data are available at: https://github.com/Veyasan1/STA3014-Paper1"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Urban spaces are often void of nature, whether it be grass fields, blooming flowers, or teaming and diverse natural wildlife. One simple way to tackle this is to incorporate nature within the urban landscape, such as with trees. Trees are a vital plant in the world, they take in carbon dioxide and provide oxygen, a key component in the air we breath. Urban areas are known for the excess in carbon dioxide, which only makes it seem natural that trees can live and perhaps thrive in such environments. Trees can also alter the way wind affects pedestrians below (@beptree). It should be no suprise then, that many cities around the world, including Toronto, have incorporated trees into their landscape, in busy downtown centres, to sleepy suburbs, and everywhere in between.

One way cities have incorportated trees into the city are roadside trees. Trees line many of the cities highways, arterials, collectors, and small roads. Large quantaties of trees lined against highways can act as a natural sound barrier, INSERT STATISTIC WITH SORUCE HERE. Even in cases where trees are more sporadic, they act as

# Data {#sec-data}

The data comes from OpenData Toronto, an open source portal containng datasets carried out at the municipal level (@opendata). This particular dataset is called Street Tree Data, and focuses on city-owned trees located on roads. This information would be most likely used by city planners and road maintenance, ensuring snow plows and street dusters don't impede on any trees that line the roads. It is important to note that there may be some privately owned trees listed among the municipal trees, as they may be of interest with regards to road maintenance or other city services.

Street Tree Data contains around 32000 observations, each observation being an individual tree. Each observation has a general id for data analysis, as well as an structure id, telling us if the tree is part of an existing structure or building. Next is the location of the tree, using the nearest parcel delivery address, nearest cross-junction, and which ward it is located in. Street Tree contains both the botanical name and the common english name, however we have decided to use only the botanical name to avoid confusion between similar sounding trees. Common english names will be provided for species of key interest. Finally, there is a measurement called dbh_trunk. This is a standard method for measuring the diameter of a tree. The method asks record takeres to read the circumference of the tree from 1.3m off the ground it is rooted in. This measurement has been known to be vaguage, especially when it comes to wawrped or tipping trees, making the exact measurements made on each tree inaccurate at best.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-topspecies
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

top_species_count <- analysis_data %>%
  count(botanical_name, sort = TRUE) %>%
  head(10)

top_species_count |> 
  ggplot(aes(y = botanical_name, x = n)) +
  geom_col() +
  theme_minimal() +
  labs(x = "Planted Trees",
       y = NULL)
```
Talk way more about it.

```{r}
#| label: fig-ward
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

ward_count <- analysis_data %>%
  count(ward, sort = TRUE) %>%
  head(10)

ward_count |> 
  ggplot(aes(x = ward, y = n)) +
  geom_col() +
  theme_minimal() +
  labs(x = "Ward",
       y = "Number of trees planted")
```

```{r}
#| label: fig-speciesward
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

top_species_ward_count <- analysis_data %>%
  filter(botanical_name == "Acer platanoides") %>%
  count(ward, sort = TRUE) %>%
  head(10)

top_species_ward_count |> 
  ggplot(aes(x = ward, y = n)) +
  geom_col() +
  theme_minimal() +
  labs(x = "Ward",
       y = "Number of trees planted")
```
```{r}
#| label: fig-trunk
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

max_dbh <- analysis_data %>%
  filter(dbh_trunk == 815)

max_dbh

analysis_data |> 
  ggplot(mapping = aes(x = dbh_trunk)) +
  geom_histogram() +
  theme_minimal() +
  labs(x = "DBH Diameter",
       y = "Tree Count")
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

\newpage

# References